---
title: "lab_7"
author: "The Pham"
date: "October 23, 2017"
output: word_document
---

#1
#Complete

#2
#Complete


#3
#a and b
#Size every cluster for number of clusters = 2,3,4,5,7,9,11 in that order
```{r echo=FALSE, warning=FALSE}
#1
#Complete

#2
#Complete

#3
#a and b
#scale only for kmeans and hac. Not for dbscan

iris<-iris
# iris_copy <- scale(subset(iris, select=c(1:4)))
iris_copy <- subset(iris, select=c(1:4))
result<-list()
for (i in 1:11){
  result[[i]]<-kmeans(iris_copy,i)
}

#Size of clusters
for (i in 1:11){
assign(paste("size",i,sep=""),result[[i]]$size)
}
#size of clusters when n of clusters =2
size2
#size of clusters when n of clusters =3
size3
#size of clusters when n of clusters =4
size4
#size of clusters when n of clusters =5
size5
#size of clusters when n of clusters =7
size7
#size of clusters when n of clusters =9
size9
#size of clusters when n of clusters =11
size11

#find a, b, c, d
#assign(paste("cluster",i,sep=""),as.vector(result[[i]]$cluster))


cluster.result<-list()
for (i in 1:11){
  
  cluster.result[[i]]<-as.vector(result[[i]]$cluster)
}

a<-function(cluster){
  if (is.null(cluster)==T){
    return(1)
  }else{
a<-0
for(i in 1:nrow(iris)){
  for (j in 1:nrow(iris)){
    if (cluster[i]==cluster[j] & iris$Species[i]==iris$Species[j]){
      a<-a+1
    }
  }
}
  if(a>=1){
    a<-a-1
  }else{
    a<-a
  }
return(a)
}
}

b<-function(cluster){
  if (is.null(cluster)==T){
    return(1)
  }else{
  a<-0
  for(i in 1:nrow(iris)){
    for (j in 1:nrow(iris)){
      if (cluster[i]!=cluster[j] & iris$Species[i]!=iris$Species[j]){
        a<-a+1
      }
    }
  }
  if(a>=1){
    a<-a-1
  }else{
    a<-a
  }
  return(a)
}
}
c<-function(cluster){
  if (is.null(cluster)==T){
    return(1)
  }else{
  a<-0
  for(i in 1:nrow(iris)){
    for (j in 1:nrow(iris)){
      if (cluster[i]==cluster[j] & iris$Species[i]!=iris$Species[j]){
        a<-a+1
      }
    }
  }
  if(a>=1){
    a<-a-1
  }else{
    a<-a
  }
  return(a)
}
}
d<-function(cluster){
  if (is.null(cluster)==T){
    return(1)
  }else{
  a<-0
  for(i in 1:nrow(iris)){
    for (j in 1:nrow(iris)){
      if (cluster[i]!=cluster[j] & iris$Species[i]==iris$Species[j]){
        a<-a+1
      }
    }
  }
  if(a>=1){
    a<-a-1
  }else{
    a<-a
  }
  return(a)
}
}
a.list<-list()
b.list<-list()
c.list<-list()
d.list<-list()

for (i in 1:11){
  a.list[[i]]<-a(cluster.result[[i]])
  b.list[[i]]<-b(cluster.result[[i]])
  c.list[[i]]<-c(cluster.result[[i]])
  d.list[[i]]<-d(cluster.result[[i]])
}

for (i in 1:11){
  assign(
    paste("f",i,sep=""),
    (2*(a.list[[i]]/(a.list[[i]]+c.list[[i]]))*
       (a.list[[i]]/(a.list[[i]]+d.list[[i]])))/
      (a.list[[i]]/(a.list[[i]]+c.list[[i]])+a.list[[i]]/(a.list[[i]]+d.list[[i]])))
}
```

#C
#Values of the F-measures for clusters=2,3,4,5,7,9,11 in that order.
```{r echo=F}
#Values of the F measures
l<-list(f2,f3,f4,f5,f7,f9,f11)
#F measure for number of clusters = 2
f2
#F measure for number of clusters = 3
f3
#F measure for number of clusters = 4
f4
#F measure for number of clusters = 5
f5
#F measure for number of clusters = 7
f7
#F measure for number of clusters = 9
f9
#F measure for number of clusters = 11
f11
```
#k=3 gives the highest F measure
#This is the value of the highest F-measure


```{r echo=FALSE, warning=FALSE}
# c

f3
#k=3 gives the highest F measure
#The number of clusters in the target is 3, while the optimal number of clusters using this
# kmeans method is also 3, as this number for k gives us the highest F measure.
#So the algorithm arrived at the target's number of clusters.

```
#d
#The number of clusters in the target is 3, while the optimal number of clusters using this
# kmeans method is also 3, as this number for k gives us the highest F measure.
#So the algorithm arrived at the target's number of clusters.

*** 

#4
#a complete
#HAC report, including the heights of the distance and the dendrogram
```{r echo=F, warning=F}
iris_dist <- dist(iris_copy)

new<-hclust(iris_dist)
# new[[2]]
plot(new)
x<-rect.hclust(new,k=3)

writeLines(as.character(x[[1]]),"a.txt")
writeLines(as.character(x[[2]]),"b1.txt")
writeLines(as.character(x[[3]]),"b2.txt")

```
#b
#judging from the plot of dendrogram, 3 clusters seem to be a reasonable threshold.
#The reason is that the heights are close together until after the number of clusters
#gets to 4, then between 4 and 3 there is a jump. The data points also seem to cluster itself
#into 2 main groups and within the right group the data partitions itself into 2
#. Therefore, 3 clusters seem reasonable.

#c
#cluster size and composition for each cluster with kmeans for cluster 1,2,3
```{r echo=FALSE, warning=FALSE}

a1<-t(read.table("a.txt"))
b1<-t(read.table("b1.txt"))
b2<-t(read.table("b2.txt"))

hac1<-as.vector(a1)
hac2<-as.vector(b1)
hac3<-as.vector(b2)
#b
#judging from the plot of dendrogram, 3 clusters seem to be a reasonable threshold.
#The reason is that the heights are close together until after the number of clusters
#gets to 4, then between 4 and 3 there is a jump. The data points also seem to cluster itself
#into 2 main groups and within the right group the data partitions itself into 2
#. Therefore, 3 clusters seem reasonable.


#cluster number and composition for each cluster with kmeans for cluster 1,2,3
#c

x<-as.vector(result[[3]]$cluster)
kmeans<-as.factor(x)

library(car)
iris2<-cbind(iris,kmeans)

hac<-cbind(hac1,1)
hac.2<-cbind(hac2,2)
hac.3<-cbind(hac3,3)
hac<-rbind(hac,hac.2,hac.3)

hac <- hac[order(hac[,1]),]
hac.factor<-as.factor(hac[,2])
iris3<-cbind(iris2,hac.factor)
#cluster number and composition for each cluster with kmeans for cluster 1,2,3
kmeans1<-which(x==1)
```
#composition and number of data points for cluster 1
```{r echo=F}
kmeans1
length(kmeans1)
```
#composition and number of data points for cluster 2
```{r echo=F}
#composition and number of data points for cluster 2
kmeans2<-which(x==2)
kmeans2
length(kmeans2)
```
#composition and number of data points for cluster 3
```{r echo=F}
kmeans3<-which(x==3)
kmeans3
length(kmeans3)
```


#cluster number and composition with HAC for cluster 1,2,3


#composition and number of data points for cluster 1
```{r echo=FALSE, warning=FALSE}
length(hac1)
hac1
```
#composition and number of data points for cluster 2
```{r echo=FALSE, warning=FALSE}
length(hac2)
hac2
```

#composition and number of data points for cluster 3
```{r echo=FALSE, warning=FALSE}
length(hac3)
hac3
```
#The number of data points in each cluster is fairly different between the
#2 methods for all 3 clusters.The composition of each cluster is fairly similar for both methods.
#The differences can be noted on the 2 graphs above

```{r echo=FALSE, warning=FALSE}
library(car)
scatterplot(Sepal.Length~Sepal.Width|kmeans,data=iris3,smoother=NULL,reg.line=NULL,main="   kmeans")
scatterplot(Sepal.Length~Sepal.Width|hac.factor,data=iris3,smoother=NULL,reg.line=NULL,main="   HAC")

```
#The number of data points in each cluster is fairly different between the
#2 methods for all 3 clusters.The composition of each cluster is fairly similar for both methods.
#The differences can be noted on the 2 graphs above
***
#5
#a complete
```{r echo=FALSE, warning=FALSE}
#5
#a

library(dbscan)
iris_mat<- as.matrix(iris_copy)
iris_mat_result<-list()

e <- ".2,.3,.4,.5,.6,.8,1"
writeLines(e,"esp.txt")
eps<-as.vector(t(read.table("esp.txt",sep=",")))
for (i in 1:length(eps)) {
  iris_mat_result[[i]] <- dbscan(iris_mat, eps = eps[i])
}
```
#b
```{r echo=FALSE, warning=FALSE}
#b
#Size of clusters
#Count number of clusters for each differing epsilon value
null<-numeric()
for (i in 1:7){
  null[i]<-length(which(iris_mat_result[[i]]$cluster==0))
}
#which(null==0)
#For epsilon =1, there is no noise data point
n<-numeric()
for (i in 1:7){
  n[i]<-length(unique(iris_mat_result[[i]]$cluster))
}
#Number of clusters for epsilon = 0.2, 0.3, 0.4, 0.5, 0.6, 0.8 or 1
#n-1
#Note: for epsilon=1, the number of clusters is 2 instead of 1
#size of clusters for differing epsilon values
```
#size of clusters for differing epsilon values
#epsilon = .2
```{r echo=FALSE}
#epsilon = .2
length(which(iris_mat_result[[1]]$cluster==1))
length(which(iris_mat_result[[1]]$cluster==2))
```
#epsilon = .3
```{r echo=F}
#epsilon = .3
length(which(iris_mat_result[[2]]$cluster==1))
length(which(iris_mat_result[[2]]$cluster==2))
length(which(iris_mat_result[[2]]$cluster==3))
```
#epsilon = .4
```{r echo=F}
#epsilon = .4
length(which(iris_mat_result[[3]]$cluster==1))
length(which(iris_mat_result[[3]]$cluster==2))
length(which(iris_mat_result[[3]]$cluster==3))
length(which(iris_mat_result[[3]]$cluster==4))
```
#epsilon = .5
```{r echo=F}
#epsilon = .5
length(which(iris_mat_result[[4]]$cluster==1))
length(which(iris_mat_result[[4]]$cluster==2))
```
#epsilon = .6
```{r echo=F}
#epsilon = .6
length(which(iris_mat_result[[5]]$cluster==1))
length(which(iris_mat_result[[5]]$cluster==2))
```
#epsilon = .8
```{r echo=F}
#epsilon = .8
length(which(iris_mat_result[[6]]$cluster==1))
length(which(iris_mat_result[[6]]$cluster==2))
```
#epsilon = 1
```{r echo=F}
#epsilon = 1
length(which(iris_mat_result[[7]]$cluster==1))
length(which(iris_mat_result[[7]]$cluster==2))
```
```{r echo=F, warning=F}
cluster.result2<-list()
for (i in 1:7){
  cluster.result2[[i]]<-as.numeric(iris_mat_result[[i]]$cluster)
}

a.list2<-list()
b.list2<-list()
c.list2<-list()
d.list2<-list()
for (i in 1:7){
  a.list2[[i]]<-a(cluster.result2[[i]])
  b.list2[[i]]<-b(cluster.result2[[i]])
  c.list2[[i]]<-c(cluster.result2[[i]])
  d.list2[[i]]<-d(cluster.result2[[i]])
}

for (i in 1:7){
  assign(
    paste("f_mat",i,sep=""),
    (2*(a.list2[[i]]/(a.list2[[i]]+c.list2[[i]]))*
       (a.list2[[i]]/(a.list2[[i]]+d.list2[[i]])))/
      (a.list2[[i]]/(a.list2[[i]]+c.list2[[i]])+a.list2[[i]]/(a.list2[[i]]+d.list2[[i]])))
}
```
#F-measure for epsilon = .2, .3, .4, .5, .6, .8, 1 in that order. The last value is the highest F-measure

```{r echo=F}

#The F-measures
#F-measure for epsilon = .2
f_mat1
#F-measure for epsilon = .3
f_mat2
#F-measure for epsilon = .4
f_mat3
#F-measure for epsilon = .5
f_mat4
#F-measure for epsilon = .6
f_mat5
#F-measure for epsilon = .8
f_mat6
#F-measure for epsilon = 1
f_mat7
#c
l2<-list(f_mat1,f_mat2,f_mat3,f_mat4,f_mat5,f_mat6,f_mat7)
#which.max(l2)
f_mat7
#F-measure is greatest for epsilon = 1
```
#c
#F-measure is greatest for epsilon = 1

#d
#When epsilon =1, there is no more noise in the clustering. It appears that as epsilon increases,
#the amount of noise data points decreases. The F-measure is also the greatest when epsilon=1.

```{r echo=F, warning=F, results='hide'}


#e
#cluster composition for each cluster with kmeans
kmeans1<-which(x==1)

#composition and number of data points for cluster 1
kmeans1
length(kmeans1)
#composition and number of data points for cluster 2
kmeans2<-which(x==2)
kmeans2
length(kmeans2)
#composition and number of data points for cluster 3
kmeans3<-which(x==3)
kmeans3
length(kmeans3)

#cluster composition with HAC
#composition and number of data points for cluster 1

length(hac1)
hac1
#composition and number of data points for cluster 2

length(hac2)
hac2
#composition and number of data points for cluster 3

length(hac3)
hac3
```


#Cluster composition with dbscan

#There are only 2 clusters
#Size and composition of data points for cluster 1 and 2
#Size and composition of data points for cluster 1
```{r echo=F, warning=F}
which(iris_mat_result[[7]]$cluster==1)
length(which(iris_mat_result[[7]]$cluster==1))
```
#Size and composition of data points for cluster 2
```{r echo=FALSE, warning=FALSE}
which(iris_mat_result[[7]]$cluster==2)
length(which(iris_mat_result[[7]]$cluster==2))
```
#dbscan method produces a clustering with only 2 clusters, different from kmeans or hac method.
#The number of data points in each cluster is fairly different amongst the
#3 methods. It is rather difficult to compare the composition of clusters in dbscan method with the
#other two methods because they don't have the same number of clusters.

#6
```{r echo=F, warning=F, error=F}
#6
swiss<-swiss
swiss_copy <- scale(swiss)
swiss_dist <- dist(swiss_copy)
fit<-hclust(swiss_dist)
plot(fit)
x2<-rect.hclust(fit,k=2)
writeLines(as.character(x2[[1]]),"catholic.txt")
writeLines(as.character(x2[[2]]),"protestant.txt")
hacswiss1<-as.vector(t(read.table("catholic.txt")))
hacswiss2<-as.vector(t(read.table("protestant.txt")))

hacswiss<-cbind(hacswiss1,1)
hacswiss.2<-cbind(hacswiss2,2)
hacswiss<-rbind(hacswiss,hacswiss.2)
hacswiss <- hacswiss[order(hacswiss[,1]),]
hac.factor<-as.factor(hacswiss[,2])
swiss<-cbind(swiss,hac.factor)
```
```{r echo=F, warning=F}
library(car)
scatterplot(Catholic~Examination|hac.factor,data=swiss,smoother=NULL,reg.line=NULL,main="HAC")
```

#Catholic-predominant cities
```{r echo=F}

rownames(swiss[hacswiss1,])
```
#Protestant-predominant cities
```{r echo=F}
#Protestant-predominant cities
rownames(swiss[hacswiss2,])

```

